---
title: "Fundamentals of Data Visualization"
output: html_notebook
author: "Marvin von Rappard"
---


```{r}
data(mtcars)
data(mpg)
library(ggplot2)
library(tidyverse)
library("ggpubr")
```

## Datasets

### mpg
```{r}
mpg
```

### mtcars
```{r}
mtcars
```


## LO1: Visualization basics, chart types

### Continuous variables

```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point() +
  labs(x = "Engine displacement (in litres)", y = "City miles per gallon (mpg)", title = "Comparison between engine displacement and miles per gallon")
```
#### Correlation between the variables

```{r}
cor(mpg$cty, mpg$displ, method="pearson")
```


```{r}
mtcars %>%
  ggplot(aes(qsec, mpg)) +
  geom_point() + 
  labs(x = "1/4 mile time", y = "Miles per gallon (mpg)", title = "1/4 mile time compared to miles per gallon (mpg)")
```

#### Correlation between the variables
```{r}
cor(mtcars$mpg, mtcars$qsec, method="pearson")
```

### Categorial Variables

```{r}
mpg_cyl <- mpg
mpg_cyl$cyl <- mpg_cyl$cyl %>% factor(levels = c(4, 5, 6, 8))

mpg_cyl %>%
  group_by(cyl) %>%
  summarize(avg_mpg = mean(cty)) %>%
  ggplot(aes(cyl, avg_mpg)) +
  geom_col() +
  labs(x = "Number of Cylinders", y = "Average city miles per gallon (mpg)", title = "Average Miles per gallon for each number of cylinders")
```

#### Correlation between the variables
```{r}
cor(mpg$cty, mpg$cyl, method="spearman")
```


```{r}
mtcars %>%
  group_by(gear) %>%
  summarize(avg_mpg = mean(mpg)) %>%
  ggplot(aes(x = gear, y = avg_mpg)) +
  geom_col() +
  labs(x = "Number of Gears", y = "Average miles per gallon (mpg)", title = "Average miles per gallon for each number of gears of a car")
```

#### Correlation between the variables
```{r}
cor(mtcars$mpg, mtcars$gear, method="spearman")
```


```{r}
mtcars_am <- mtcars
mtcars_am$am <- mtcars_am$am %>% factor(levels=c(0, 1), labels=c("Manual", "Automatic"))

mtcars_am %>%
  group_by(am) %>%
  summarize(avg_mpg = mean(mpg)) %>%
  ggplot(aes(am, avg_mpg)) +
  geom_col() + 
  labs(x = "Transmission type", y = "Average miles per gallon (mpg)", title = "Average miles per gallon by transmission type")
```

```{r}
mtcars_am %>%
  ggplot(aes(x = am, y = mpg)) + 
  geom_boxplot() +
  labs(x = "Transmission type", y = "Miles per gallon (mpg)", title = "Spread of miles per gallon by transmission type")
```

## LO2: Visual Perception

### Colors

```{r}
ggplot(mpg, aes(displ, cty, color=cyl)) +
  geom_point() +
  labs(x = "Engine displacement (in litres)", y = "City miles per gallon (mpg)", title = "Comparison between engine displacement and miles per gallon")
```

```{r}
ggplot(mpg_cyl, aes(displ, cty, color=cyl)) +
  geom_point() +
  scale_color_hue(l=40, c=35) +
  labs(x = "Engine displacement (in litres)", y = "City miles per gallon (mpg)", title = "Comparison between engine displacement and miles per gallon")
```

### Shapes
```{r}
mpg_manu <- mpg %>%
  group_by(manufacturer) %>%
  summarize(avg_mpg = mean(cty)) %>%
  arrange(avg_mpg) %>%
  slice(1:5)

pie(mpg_manu$avg_mpg, labels= mpg_manu$manufacturer, col=rainbow(length(mpg_manu$manufacturer)), main="Average city miles per gallon (mpg) per manufacturer")
```
```{r}
mpg_manu %>% ggplot(aes(manufacturer, avg_mpg, fill = manufacturer)) +
  geom_col() +
  labs(x = "Manufacturer", y = "Average city miles per gallon (mpg)", title = "Average city miles per gallon (mpg) per manufacturer")
```



### Scale

```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point() +
  scale_y_log10() +
  labs(x = "Engine displacement (in litres)", y = "City miles per gallon (mpg, log10)", title = "Comparison between engine displacement and miles per gallon")
```
## LO4: Data Preprocessing

### Adding helper columns with preprocessed information

```{r}
mpg_enhanced <- mpg
mpg_enhanced$mpg_combined <- rowMeans(mpg[,c(8, 9)])

mpg_enhanced
```

```{r}
mpg_enhanced %>% ggplot(aes(displ, mpg_combined)) +
  geom_point() +
  labs(x = "Engine Displacement (in liters)", y = "Combined miles per gallon (mpg)", title = "Engine dsiplacement vs combined miles per gallon")
```

```{r}
mpg_enhanced$transmission_type <- substr(mpg$trans, 1, 4)
mpg_enhanced
```

```{r}
mpg_enhanced %>% 
  ggplot(aes(transmission_type, mpg_combined)) +
  geom_boxplot() +
  labs(x = "Transmission type", y = "Average combined miles per gallon", title= "Distribution of miles per gallon per type of transmission")
```

### Factorizing categorical variables

```{r}
mpg_enhanced$transmission_type <- mpg_enhanced$transmission_type %>% factor(levels = c("manu", "auto"), labels = c("Manual", "Automaitc"))
mpg_enhanced$cyl <- mpg_enhanced$cyl %>% factor(levels = c(4, 5, 6, 8), labels = c("4 Cylinder", "5 Cylinder", "6 Cylinder", "8 Cylinder"))
mpg_enhanced$drv <- mpg_enhanced$drv %>% factor(levels = c("r", "f", "4"), labels = c("Rear wheel drive", "Front wheel drive", "All wheel drive"))
mpg_enhanced
```

```{r}
mpg_enhanced %>% 
  group_by(drv) %>%
  summarize(avg_mpg = mean(mpg_combined)) %>%
  ggplot(aes(drv, avg_mpg)) +
  geom_col() +
  labs(x = "Drive train type", y = "Average miles per gallon", title = "Average miles per gallon per drive train type")
```

```{r}
mpg_enhanced %>% 
  group_by(cyl) %>%
  summarize(avg_mpg = mean(mpg_combined)) %>%
  ggplot(aes(cyl, avg_mpg)) +
  geom_col() +
  labs(x = "Number of cylinders", y = "Average miles per gallon", title = "Average miles per gallon per number of cylinders in an engine")
```


